Community-Conditioned Language Models

### 2021-05-14 paper/paper.tex

Computing this distance is known as the orthogonal Procrustes problem~\citep{Gower2004}. Its
solution is
\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). The
singular values \(\sigma_i\) (diagonal elements of \(Σ\)), say how much
correlation there is between the directions \(U_i\) and \(V_i\).

Further, we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
is, the largest singular value \(σ_0\) corresponds to the principal
directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
principal direction, etc.

The \(d(L,S)\) metric varies between \(0\) (perfect independence) and
\(n\) (perfect correlation, obtained for example if \(L=S\)), where
$n$ is the number of community embeddings considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have computed experimentally \(d(L,S)\) for 
\(L\) and \(S\) chosen at random and 16 dimensions. We find
a mean of \num{73.83} and a standard deviation of \num{1.89}.

Thus \(d(L,S)\) below \(μ-3σ\) (\num{68.15}) can be safely considered to indicate a \bn{above  \(μ+3σ\) (\num{79.50})?}
statistically significant correlation between \(L\) and
\(S\). However, below, we will also give a sense of \emph{how} the
correlation is manifested by analysis of the two principal components
of correlation in the linguistic embeddings, \(U_0\) and \(U_1\).

>>>>

Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). Each
singular \(\sigma_i\) (elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

Further, we can apply a permutation to \(U\),\(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). That
is, the largest singular value \(σ_0\) then corresponds to the principal
directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
principal direction, etc.

The \(d(L,S)\) metric range from \(0\) (corresponding to perfect
independence) to \(n\) (corresponding to perfect correlation, obtained
for example if \(L=S\)), where $n=510$ is the number of communities
considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have computed experimentally \(d(L,S)\) for 
\(L\) and \(S\) chosen as random 16 dimensions. We find
a mean of TODO and a standard deviation of TODO.

TODO: !!! we should take a 16 × 510 and a 300 × 510 matrix to test!!!

Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there \bn{above  \(μ+3σ\) (\num{79.50})?}
is statistically significant correlation between \(L\) and
\(S\). However, below, we will also give a sense of \emph{how} the
correlation is manifested by analysis of the two principal components
of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
do so we will plot the projection of embeddings on along \(U_0\) and
\(U_1\), together with the corresponding singular values.
